---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dsn

<!-- badges: start -->
[![R-CMD-check](https://github.com/hypertidy/dsn/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/hypertidy/dsn/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of dsn is to provide simple helpers for GDAL data source name prefixes and related string handling. 

## Installation

You can install the development version of dsn from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hypertidy/dsn")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(dsn)
vsicurl("https://netcdf-r-us.org/f.nc")

driver("somefile.h5", "HDF5")

unvsicurl("/vsicurl/https://netcdf-r-us.org/f.nc")

unprefix("NETCDF:/u/user/somefile.nc")
```



## Code of Conduct
  
Please note that the dsn project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.
